# Recitation 11

* For loops

Now we'll practice with iteration.

Load any necessary packages in your first chunk of code.

```{r, include=FALSE}
library(tidyverse)
```

Please work out the following exercises sequentially. There may be more than one way to code the exercises.

## Part I: Iteration

1. `x` is a 4x3 data frame of 12 randomly generated numbers from a Uniform(0, 1) distribution. Write a for loop that returns the standard deviation of each column of `x` in the following format:

[1] "1 : 0.203624272520283"

[1] "2 : 0.297398365605958"

[1] "3 : 0.370973895270821"

Do not hard code the indices, 1, 2, 3. 

```{r}
x <- as.data.frame(matrix(runif(12), ncol = 3))

for (i in 1:ncol(x)) {
  print(str_c(i, ": ", sd(x[,i])))
}
```

2. Write a nested for loop (a for loop inside a for loop) that returns an output that looks something like this:

[1] "Element x[1, 1] is 0.293645224068314" 

[1] "Element x[1, 2] is 0.691126650897786"

[1] "Element x[1, 3] is 0.803303817752749" 

[1] "Element x[2, 1] is 0.385301414178684" 

[1] "Element x[2, 2] is 0.128541933139786"

[1] "Element x[2, 3] is 0.131027936702594" 

[1] "Element x[3, 1] is 0.43460671370849" 

[1] "Element x[3, 2] is 0.874724538065493" 

[1] "Element x[3, 3] is 0.0957115420605987" 

[1] "Element x[4, 1] is 0.826694968156517" 

[1] "Element x[4, 2] is 0.826714764349163" 

[1] "Element x[4, 3] is 0.900152393151075" 

```{r }
for (i in 1:nrow(x)) {
  for (j in 1:ncol(x)) {
    print(str_c("Element x[", i, ", ", j, "] is ", x[i, j]))
  }
}
```

3. *CHALLENGE* Write a function called `remove.let` that goes through each word of a string and removes the first letter. Use the given strings to test your function once it is written. `stringr` functions will be helpful here and you _must_ use iteration.

My general roadmap:

* Convert the string into a vector of individual words
* Go through each word in the vector and remove the first letter, using iteration
* Put the individual words back into one string
* Return the string

```{r}
remove.let <- function(s) {
  
  s_sep <- str_split(s, pattern = " ")[[1]]
  
  output <- vector("character", length = length(s_sep))
  
  for (i in 1:length(s_sep)) {
    output[i] <- str_sub(s_sep[i], 2, -1)
  }
  
  return(str_c(output, collapse = " "))
  
}
  
# Check
my_string <- "We love recitation"
my_string2 <- "My favorite letter is R"

remove.let(my_string) # Answer should be: "e ove ecitation
remove.let(my_string2) # Answer should be: "y avorite etter s "
```
