# Recitation 1 

## Objectives
- Download `R` and `RStudio`
- Create R project
- Get familiar with `.R` script and `.Rmd` notebook
- Variables
- R Objects


## Download Softwares

### Installing R

#### Windows

1. Go to the [CRAN R Project website](https://cloud.r-project.org/).
2. Click on the link to download R for Windows.
3. Click on the "base" link.
4. Click on the "Download R for Windows" link to download the installer.
5. Once the installer is downloaded, open it and follow the on-screen instructions to complete the installation.

#### macOS

1. Go to the [CRAN R Project website](https://cloud.r-project.org/).
2. Click on the link to download R for macOS.
3. Click on the `.pkg` file link to download the installer.
4. Once the installer is downloaded, open it and follow the on-screen instructions to complete the installation.


### Installing RStudio

#### Windows

1. Go to the [RStudio website](https://posit.co/download/rstudio-desktop/).
2. Under the "Installers for Supported Platforms" section, click on the link for Windows to download the installer.
3. Once the installer is downloaded, open it and follow the on-screen instructions to complete the installation.

#### macOS

1. Go to the [RStudio website](https://posit.co/download/rstudio-desktop/).
2. Under the "Installers for Supported Platforms" section, click on the link for macOS to download the installer.
3. Once the installer is downloaded, open it and follow the on-screen instructions to complete the installation.


## Create Your First Project
1. **Open RStudio:** Launch RStudio after installation.
2. **Create a New Project:**
   - Go to `File` > `New Project`.
   - Choose `New Directory`, then `New Project`.
   - Select a folder on your computer where you want to save your project.
   - Click `Create Project`.

This sets up a new R project where you can start writing and organizing your R scripts and data files.

## Different Files
### `.R` Script
`.R` files are simple text files containing R code. They are used for writing and running R scripts, which can include data analysis, data manipulation, and plotting commands.

Creating an .R Script:
1. In RStudio, go to `File` > `New File` > `R Script`.

2. Write your R code in the script editor.

3. Save the file with a `.R` extension.

### `.Rmd` Rmarkdown
`.Rmd` (R Markdown) files are used for creating dynamic documents that combine code, text, and visualizations. They are ideal for reports, presentations, and reproducible research.

Creating an `.Rmd` Notebook:
1. In RStudio, go to `File` > `New File` > `R Markdown`.
2. Choose the type of output (e.g., HTML, PDF, Word).
3. Write your code and narrative text in the R Markdown editor.
4. Use R code chunks (```{r}) to embed R code.
5. Click the `Knit` button to generate the final document.

**Keyboard Shortcuts: **


|     Operation     |    Windows     |     MacOS     |
| :---------------: | :------------: | :-----------: |
| create code chunk | `Ctrl + Alt + I` | `Cmd + Opt + I` |
|  run code chunk   |  `Ctrl + Enter`  | `Cmd + Return`  |

## Variables
### Coming up with variable names
Use concise and meaningful names for your variables. Object names must start with a letter and can only contain **letters**, **numbers**, **_** and **.** 
Base R uses dots in function names (ex: `contrib.url()`) and class names (ex: `data.frame`). I recommend the snake_case where you can separate words with _.

### Numeric

Numeric variables store numbers and can be either integers or doubles (floating-point numbers).

```{r}
x <- 42  # integer
y = 3.14  # double
```

```{r}
# addition 
x + y

# substraction
x - y

# multiplication
x * y

# division
x / y

# mod
x %% 9
```

### String

String variables store text data and are created by enclosing the text in quotes.

```{r}
name <- "John Doe"
message <- "Hello, world!"
```

### Logical 

Logical variable store data as value which is `TRUE` or `FALSE`.
```{r}
correct = TRUE
correct = T
wrong = FALSE
wrong = F
```

## R Objects

### Tips
You could know **everything** about R programming with the following command: 
```{r eval=FALSE}
# help document for function
?mean

# help document for package
??MASS
```

### Vector

Vectors are one-dimensional arrays that can hold numeric, string, or logical data. They are created using the `c()` function.

```{r}
num_vector <- c(1, 2, 3, 4, 5)
str_vector <- c("apple", "banana", "cherry")
```

### List

Lists are versatile data structures that can hold different types of data, including other lists.

```{r}
# definition
my_list <- list(name = "John", age = 30, scores = c(90, 85, 88))
# slice
my_list[1:2]
# pull out element
my_list[[1]]
my_list[["name"]]
my_list$name
```

### Factor

Factors are used to represent categorical data and are useful for statistical modeling.

```{r}
colors <- factor(c("red", "blue", "green", "blue", "red"))
colors
```

### Matrix

Matrices are two-dimensional arrays that hold numeric data. They are created using the `matrix()` function.

```{r}
my_matrix <- matrix(1:9, nrow = 3, ncol = 3)
my_matrix
```

### DataFrame

DataFrames are table-like structures that can hold different types of data in columns, making them ideal for datasets.
```{r eval=FALSE}
install.packages("nycflights13")
```

```{r}
library(nycflights13)

flights
```
```{r}
# To view the structure of the df
str(flights)

# Summary stats of the numeric variables
summary(flights)

# Finding the dimensions of the df
dim(flights)

# Mean calculation
mean(flights$distance)
# Mean calculation with NA clean
mean(flights$distance, na.rm = TRUE)

# Max value
max(flights$air_time)
# Max value with NA clean
max(flights$air_time, na.rm = TRUE)

# Unique value
unique(flights$dest)
```
With these basics, you should be able to get started with your first R project, create different types of files, and understand the fundamental data structures in R.

## Some Spacing Conventions
* Include spaces before and after +, -, *, /
* Don't put spaces around ^ (e.g., 3^2)
* Don't put spaces around : or $
* Include spaces after commas (e.g., c(0, 1, 2))

## Other Resource
This page touches on some of the conventions we discussed (if you're interested): http://adv-r.had.co.nz/Style.html

More on Basics and Workflow: https://stat545.com/r-basics.html



