# Recitation 2

__Today you learned...__

* Pipe operators
* Importing data

Now we'll practice these concepts (and some stuff from Recitation 1).

The first step, which we did together, was setting up a project for this recitation. 

Make sure when you go to your `Files` pane, or the project icon in the top right, to see you are working in the right project/directory.

Next, load any necessary packages in a first chunk of code and name this chunk "setup" with `include = FALSE`. You can learn more about chunk options at this link: https://rmarkdown.rstudio.com/lesson-3.html

```{r setup, include=FALSE}
library(tidyverse)
```

Note: There may be more than one way to code the exercises.

## Part I

1. Create a variable called `numbers` that contains 1 through 10 as characters and two
`NA`s (they may be anywhere in the vector).

```{r p1q1}
numbers <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", rep(NA, 2))
```

2. Omit any `NA`’s from `numbers` and convert the character variable to a numeric one. Save
the new variable as `numbers`. Use the pipe operator: `%>%` or `|>`. 

Mac shortcut: `Shift + Command + M`

PC shortcut: `Ctrl + Shift + M`

```{r p1q2}
numbers <- numbers |> 
  na.omit() |>
  as.numeric()
```

3. Find the mean of the first 5 elements without typing out the numbers. 

```{r p1q3}
mean(numbers[1:5])
```

4. Find the mean of the first, third and fifth elements without typing out the numbers.

```{r p1q4}
mean(c(numbers[1], numbers[3], numbers[5]))
```

## Part II

__Nice! You completed Part I.__ 

Now let's load in the data for part II.

```{r data, eval=FALSE}
data_vacc <- read_csv("path/to/data_vacc.csv")
```

```{r data-true, include=FALSE}
data_vacc <- read_csv("./dataset/data_vacc.csv")
```


1. Data type:

(a) Write out the type for each variable.

* __names__: `character`
* __gender__: `character`
* __age__: `character`
* __height__: `double`
* __weight__: `double`
* __vaccinated__: `character`

(b) The `age` variable should be a numeric variable and `vaccinated variable
should be a logical variable, but there are some typos preventing this. 

Fix the typos and then use `as.numeric()` and `as.logical()` to convert them.

```{r p2q1b}
data_vacc$age[1] <- 57
data_vacc$age <- as.numeric(data_vacc$age)

data_vacc$vaccinated[3] <- FALSE
data_vacc$vaccinated <- as.logical(data_vacc$vaccinated)
```

2. Summarize the data using `skimr`

```{r p2q2}
skimr::skim(data_vacc)
```

3. Using the pipe operator, omit any `NA`s from the `weight` variable and find the mean.

```{r p2q3}
data_vacc$weight %>%
 na.omit() %>%
 mean()
```

4. Challenge! Take the subset of the data for those who are vaccinated and find their
mean age using `[]`, `$` and `%>%`. 

```{r p2q4}
data_vacc[data_vacc$vaccinated == TRUE, ]$age %>%
 mean()
```


## Part III

For this part, install the `MASS` package and call it for use in this session. 

Pull out the `birthwt` data frame and assign it to an object in your personal environment. 
```{r install-MASS, eval=FALSE}
install.packages("MASS")
```

```{r gapminder-set-up, message=FALSE}
library(MASS)
bwt = birthwt
View(bwt)
```

Explore around! 

Make and answer 2-3 (using pipes and functions you've come across so far) questions of your interest and share your approach with the class! 
