# Recitation 2

__Today you learned...__

* Pipe operators
* Importing data

Now we'll practice these concepts (and some stuff from Recitation 1).

The first step, which we did together, was setting up a project for this recitation. 

Make sure when you go to your `Files` pane, or the project icon in the top right, to see you are working in the right project/directory.

Next, load any necessary packages in a first chunk of code and name this chunk "setup" with `include = FALSE`. You can learn more about chunk options at this link: https://rmarkdown.rstudio.com/lesson-3.html

```{r}

```

Note: There may be more than one way to code the exercises.

## Part I

1. Create a variable called `numbers` that contains 1 through 10 `as characters` and two
`NA`s (they may be anywhere in the vector).

```{r p1q1}

```

2. Omit any `NA`’s from `numbers` and convert the character variable to a numeric one. Save
the new variable as `numbers`. Use the pipe operator: `%>%` or `|>`. 

Mac shortcut: `Shift + Command + M`

PC shortcut: `Ctrl + Shift + M`

```{r p1q2}

```

3. Find the mean of the first 5 elements without typing out the numbers. 

```{r p1q3}

```

4. Find the mean of the first, third and fifth elements without typing out the numbers.

```{r p1q4}

```

## Part II

__Nice! You completed Part I.__ 

Now let's load in the data for part II.

```{r data}

```

1. Data type:

(a) Write out the type for each variable.

* __names__: 
* __gender__: 
* __age__: 
* __height__: 
* __weight__: 
* __vaccinated__: 

(b) The `age` variable should be a numeric variable and `vaccinated variable
should be a logical variable, but there are some typos preventing this. 

Fix the typos and then use `as.numeric()` and `as.logical()` to convert them.

```{r p2q1b}

```

2. Summarize the data using `skimr`

```{r p2q2}

```

3. Using the pipe operator, omit any `NA`s from the `weight` variable and find the mean.

```{r p2q3}

```

4. Challenge! Take the subset of the data for those who are vaccinated and find their
mean age using `[]`, `$` and `%>%`. 

```{r p2q4}

```


## Part III

For this part, install the `MASS` package and call it for use in this session. 

Pull out the `birthwt` data frame and assign it to an object in your personal environment. 

```{r gapminder-set-up}

```

Explore around! Make and answer 2-3 (using pipes and functions you've come across so far) questions of your interest and share your approach with the class! 
