# Recitation 3

__Today you learned...__

* dplyr functions: _select_, _filter_, _mutate_, _arrange_, _rename_

Now we'll practice using these functions and refresh ourselves on a few basics from last week.

Load any necessary packages in your first chunk of code (name this chunk `setup` with `include = FALSE`).

```{r setup, include=FALSE}
detach(package:MASS, unload = T)
library(tidyverse)
```

Now, with your partner, please work out the following exercises sequentially. There may be more than one way to code the exercises.

## Part I

0. Print your working directory to show that you are located in an appropriate place on your device.

```{r p1q0}
getwd()
```

1. (Review) Create a character vector called numbers, with the numbers 1 to 5 and one NA. Then, use the pipe operator to convert this to a numeric vector, omit any `NA`’s, and find the mean. Do not save the mean (i.e. do not use "<-" to save the operations on the numbers variable).

```{r p1q1}
numbers <- c("1", "2", "3", "4", "5", NA)

numbers %>% 
  as.numeric() %>% 
  na.omit() %>% 
  mean()
```

2. (Review) Save the `diamonds.csv` file from Courseworks into a `data` folder in your current working directory. Load in this data frame and save it as `dat_diamond`.

```{r p1q2}
dat_diamond <- read_csv("dataset/diamonds.csv")
```

3. Remove the first column using one of the functions from class. Make sure this column is gone when you view the data frame.

```{r p1q3}
dat_diamond <- dat_diamond %>% select(-1)

dat_diamond <- dat_diamond %>% select(carat:z)
```

4. Find three unique ways to select the carat column (do not save the column to the environment).

```{r p1q4}
# Method 1
dat_diamond %>% select("carat")
  
# Method 2
dat_diamond %>% select(carat)
  
# Method 3
dat_diamond %>% select(1)

# Method 4
dat_diamond[, 1]
```

5. Find the mean of the carat column (do not save).

```{r p1q5}
mean(dat_diamond$carat)

# OR

dat_diamond %>% 
  pull(carat) %>% 
  mean(na.rm = TRUE)

# OR

dat_diamond %>% 
  select(carat) %>% 
  summarise(mean_carat = mean(carat))

# OR

dat_diamond %>% 
  summarise(mean_carat = mean(carat))
```

6. Create a variable called `data_prem` that only contains diamonds with a Premium cut. Also, rename the "col" variable to "color" and create a new column called "expensive" that is `TRUE` if the price is more than $5,000 and `FALSE` otherwise.

```{r p1q6}
dat_prem <- dat_diamond %>% 
  filter(cut == "Premium") %>% 
  rename(color = col) %>% 
  mutate(expensive = if_else(price > 5000, TRUE, FALSE))

# OR

data_prem = dat_diamond %>% 
  filter(cut == "Premium") %>% 
  rename(color = col) %>% 
  mutate(expensive = case_when(price > 5000 ~ TRUE, 
                               price <= 5000 ~ FALSE))
```

7. Take your `dat_prem` data set and arrange it by price in descending order. Use the function `head(5)` to print only the first 5 elements.

```{r p1q7}
dat_prem %>% 
  arrange(desc(price)) %>% 
  head(5)
```

## Part II (bonus)

1. Find the average price of Premium diamonds with color G. [Hint: last line of code that finds the mean should be `summarize(avg_price = mean(price))`.]

```{r p2q1}
dat_prem %>% 
  filter(color == "G") %>% 
  summarize(avg_price = mean(price))
```

2. For diamonds with Good or Fair cut, create a new column called ‘discounted’ that shows the price after a 10% discount and find the mean and standard deviation of the discounted prices. [Hint: last line of code that finds the mean and sd should be `summarize(avg_discount = mean(price), sd_discount = sd(price))`.]

```{r p2q2}
dat_diamond %>% 
  filter(cut %in% c("Good", "Fair")) %>% 
  mutate(discounted = price * 0.9) %>% 
  summarize(avg_discount = mean(discounted),
            sd_discount = sd(discounted))

# OR

dat_diamond %>% 
  filter(cut == "Good" | cut == "Fair") %>% 
  mutate(discounted = 0.9*price) %>% 
  summarize(avg_discount = mean(discounted), 
            sd_discount = sd(discounted))
```

3. For diamonds with clarity VS1 and carat greater than 0.8, count the number of observations that belong to each type of cut and arrange them in increasing order. [Hint: to count the frequencies of a categorical variable like cut, use `count(cut)`. For arranging, use `arrange(n)`.]

```{r p2q3}
dat_diamond %>% 
  filter(clarity == "VS1" & carat > 0.8) %>% 
  count(cut) %>% 
  arrange(n)

# OR

dat_diamond %>% 
  filter(clarity == "VS1" & carat > 0.8) %>% 
  group_by(cut) %>% 
  summarise(counts = n())
```

### Additional notes

In class, we talked about methods to return a single column as a data frame (aka, tibble) versus methods to return a single column as a vector of values. 

To return a data frame:

```{r, results=FALSE}
# Method 1
dat_diamond %>% select(carat)

# Method 2
dat_diamond[, 1]
```

To return a vector:

```{r, results=FALSE}
# Method 1
dat_diamond %>% pull(carat)

# Method 2
dat_diamond$carat
```

In order to use some functions, like `mean()` you will need to use `pull()` or the "$". Keep this in mind for future coding!
