# Recitation 4

__Today you learned...__

* pivots: _pivot_longer()_, _pivot_wider()_
* joins: _right_join()_, _left_join()_, _inner_join()_, _full_join()_

This worksheet will help you get accustomed to using these functions to manipulate data.

Load any necessary packages in your first chunk of code.

```{r, include=FALSE}
library(tidyverse)
```

Now, with your partner, please work out the following exercises sequentially. There may be more than one way to code the exercises.

## Part I

0. Load the data you'll need for Part I by simply running the code below.

```{r }
year <- paste(2001:2020)
team <- letters[1:length(year)]
num <- sample(c(1:100), size = length(year)^2, replace = TRUE)

yr_data <- data.frame(matrix(num, nrow = length(year), ncol = length(year))) %>%
  set_names(year)

index <- 1:20

tbi_age <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_age.csv")
```

1. Add `index` to the `yr_data` data frame via binding. Call the index column in the resulting data frame `idx`.

```{r}
yr_data <- bind_cols(idx = index, yr_data)
```

2. Is the data currently in wide or long format? Justify your answer with a complete sentence.

[WIDE.]

3. Create one `year` column and put all data currently stored in columns 2001 to 2020 into a column called `points`. Save this updated data frame as `yr_data_long`.

```{r }
yr_data_long <- pivot_longer(yr_data,
                            cols = paste(2001:2020),
                            names_to = "year",
                            values_to = "points")
```

4. View the `tbi_age` data and decide what format it's in. Justify your answer.

[LONG.]

5. Transform this data so that we can see `number_est` for each age group along with `type` and `injury_mechanism`, like below. Save this data frame as `tbi_wide`. 

```{}
## # A tibble: 21 x 13
##    type  injury_mechanism  0-17   0-4<code>   5-14</code>   15-24   25-34   35-44   45-54
##    <chr> <chr>             <dbl>  <dbl>  <dbl>   <dbl>   <dbl>   <dbl>   <dbl>
##  1 Emer~ Motor Vehicle C~  47138   5464  19785  103892   71641   44108   40020
##  2 Emer~ Unintentional F~ 397190 230776 133084   96568   70210   68830   95127
##  3 Emer~ Unintentionally~ 229236  53436 120839  106679   44404   32479   30495
##  4 Emer~ Other unintenti~  55785  12007  30656   37118   22360   17541   17808
##  5 Emer~ Intentional sel~     NA     NA     NA     870     650     421     247
##  6 Emer~ Assault           24360    674   9690   65399   57213   34100   27682
##  7 Emer~ Other or no mec~  57983  19360  26022   33395   20974   16503   15962
##  8 Hosp~ Motor Vehicle C~   5830    870   2395   12925   11050    7305    8490
##  9 Hosp~ Unintentional F~   7935   4700   2270    3910    4470    5640   12010
## 10 Hosp~ Unintentionally~   1985    510    980    1070     635     610     685
## # ... with 11 more rows, and 4 more variables
```

```{r }
tbi_wide <- tbi_age %>% 
  pivot_wider(
    id_cols = c(1, 2), 
    names_from = age_group,
    values_from = number_est)
tbi_wide
```

## Part II

In this example, we have a data frame with head counts for grain feeding producers and another data frame with head counts from grass feeding producers; there is potentially overlap between the producers that have both grain and grass fed animals and we want to merge them into one data frame. Note that producer number is not unique, but it is unique within state, so we need to merge on both state and producer number together.

0. Load the data for Part II by simply running the code below. 

```{r }
grass <- tibble(
  producer = c(1, 2, 3, 1),
  state = c("NSW", "NSW", "QLD", "QLD"),
  count = c(59, 82, 37, 28)
  )

grain <-  tibble(
  producer = c(1, 1, 2, 4),
  state = c("NSW", "QLD", "NSW", "QLD"),
  count = c(192, 21, 37, 81),
  post_code = c("2680", "4700", "2650", "4715")
  )
```

1. Keep all the observations that are present in the `grass` data frame and join it with the `grain` data frame using the code below. What does the argument `by = c("producer","state")` do?

```{r}
gg_lj <- left_join(grass, grain, by = c("producer", "state"))
```

[YOUR ANSWER HERE]

2. Add the argument `suffix = c("_grass", "_grain")` to the line of code below. What does this argument do?

```{r }
gg_lj2 <- left_join(grass, grain, by = c("producer", "state"), 
                    suffix = c("_grass", "_grain"))
```

[YOUR ANSWER HERE]

3. Join the data frames by writing code that keeps all the observations that are present in `grain`. 

```{r}
gg_rj <- right_join(grass, grain, by = c("producer", "state"), 
                    suffix = c("_grass", "_grain"))

# OR

gg_rj <- left_join(grain, grass, by = c("producer", "state"), 
                    suffix = c("_grain", "_grass"))
```

4. Are there any missing data? Why does this happen?

Yes. The property of `left_join` and `right_join` caused that phenomenon. Some values occured in left(right) table don't occur in right(left) table.

5. Use a join that keeps any observations that occur in either of the two data frames.

```{r}
gg_fj <- full_join(grass, grain, by = c("producer", "state"), 
                    suffix = c("_grass", "_grain"))
```

6. Looking at this merged data frame, can you tell whether producer 3 is a grass or grain feeding producer? How?

Producer 3 belongs to grass feeding produce. Because there is no grain-feeding count of producer 3 is NA, which means there is no grain-feeding animals.

7. Use a join that keeps only observations that are present in both data frames.

```{r}
gg_ij <- inner_join(grass, grain, by = c("producer", "state"), 
                    suffix = c("_grass", "_grain"))
```

8. What are the dimensions of this merged data? Why do we see what we see?

The dimension of merged data is `r dim(gg_ij)`. We see such a result because `inner_join` function only output the data item occurring in both left and right dataframe.

## Part III (Bonus Questions)

1. Clean the `tbi` data so that it contains the following age groups: 0-4, 5-17, 18-24, 25-34, 35-44, 45-54, 55-64, 65-74, 75+, and Total. Keep the data in wide format. Save your new data frame as `tbi_clean`.

```{r}
tbi_clean <- tbi_wide %>% 
  mutate(`5-17` = `0-17` - `0-4`,
         `18-24` = Total - (`0-17` + `25-34` + `35-44` + 
                              `45-54` + `55-64` + `65-74` + `75+`) 
  ) %>% 
  select(c(type, injury_mechanism, `0-4`, `5-17`, `18-24`, `25-34`:Total))

# Check that the categories sum to the total
tbi_clean %>% 
  mutate(Total2 = rowSums(across(`0-4`:`75+`), na.rm = T)) %>% 
  select(Total, Total2)
```

2. Start with the farming data frame you created in Part II Q5. Add in the latitude and longitude associated with each post code in the data set. The code below will load in a file with lat and long info for each post code.

```{r}
pcodes <- read_csv("http://garthtarr.com/data/aus_postcode_lat_long.csv") %>% 
  rename(post_code = postcode)

gg_location <- left_join(gg_fj, pcodes, by = "post_code")
```
