# Recitation 5

__Today you learned...__

* String and factor manipulation using `stringr` and `forcats`

Now we'll practice using the functions we learned in class, and we'll also get some experience with learning about functions from a written resource rather than from an instructor!

Load any necessary packages in your first chunk of code.

```{r , include=FALSE}
library(tidyverse)
```

Now, with your partner, please work out the following exercises sequentially. There may be more than one way to code the exercises.

## Part I: Strings

1. Find the length of each of the following using a `stringr` function: 'i', 'like', 'programming', and `NA`.

```{r }
str_length(c("i", "like", "programming", NA))
```

2. Write code that returns a logical (TRUE, FALSE) vector indicating whether 'apple', 'banana', 'kiwi' and 'pineapple' each contain the letter 'a'.

```{r}
fruit <- c("apple", "banana", "kiwi", "pineapple")
str_detect(fruit, "a")
```

3. Write code that returns a logical (TRUE, FALSE) vector indicating whether 'apple', 'banana', 'kiwi' and 'pineapple' each contain any vowels, a,e,i,o,u.

```{r}
str_detect(fruit, "[aeiou]")
```

4. Write code that replaces the first vowel in the given 'fruits' variable with a '-'.

```{r}
fruits <- c("one apple", "two pears", "three bananas")
str_replace(fruits, "[aeiou]", "-")
```

5. Write code that replaces all vowels in the given 'fruits' variable with a '-'.

```{r}
str_replace_all(fruits, "[aeiou]", "-")
```

6. `letters` is a built-in character vector in R that contains all 26 letters of the alphabet. Write a single line of code to get the following output:

"a, b, c, d, e, f, g, h, i, j, k, l, m, n, o, p, q, r, s, t, u, v, w, x, y, z"

```{r}
# Method 1
str_flatten(letters, collapse = ", ")

# Method 2
str_c(letters, collapse = ", ")
```

7. Write a single line of code using a `stringr` function to get the following output:

```{r}
#>  [1] "a comes before b" "b comes before c" "c comes before d"
#>  [4] "d comes before e" "e comes before f" "f comes before g"
#>  [7] "g comes before h" "h comes before i" "i comes before j"
#> [10] "j comes before k" "k comes before l" "l comes before m"
#> [13] "m comes before n" "n comes before o" "o comes before p"
#> [16] "p comes before q" "q comes before r" "r comes before s" 
#> [19] "s comes before t" "t comes before u" "u comes before v" 
#> [22] "v comes before w" "w comes before x" "x comes before y" 
#> [25] "y comes before z"
```
Hint: What are the outputs for letters[-1] and letters[-26]?

```{r}
str_c(letters[-26], " comes before ", letters[-1])
```

8. Use the `paste0` function to get the same output as in 7. 

```{r }
paste0(letters[-26], " comes before ", letters[-1])
```

## Part II: Factors

1. The `bdays` variable contains some months of the year. Apply the sort function to it. What does it do? Does it sort in a meaningful way?

```{r }
bdays <- c("Dec", "Apr", "Jan", "Mar")
sort(bdays)
```

2. Create a vector called `month_levels` that lists all months of the year in order (i.e., Jan, Feb, ...). 

```{r }
month_levels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                   "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
```

3. Create a factor variable using the `bdays` variable with `month_levels` as the levels. Now sort your new variable. What has changed?

```{r }
# Method 1: base R method
bdays_factor <- factor(bdays, levels = month_levels)

# Method 2: forcats method
bdays_factor2 <- bdays %>% as_factor()
levels(bdays_factor2) <- month_levels

sort(bdays_factor)
```

4. `f` is a vector containing 1000 samples from the alphabet following a Poisson distribution with lambda = 10. Count how many of each letter was sampled. 

```{r}
f <- factor(sample(letters)[rpois(1000, 10)])
fct_count(f)
```

5. Count how many of each letter was sampled and show them in decreasing order. 

```{r}
f %>% fct_count() %>% arrange(desc(n))

fct_count(f, sort = T)
```

6. Use the `unique()` function in base R to show the unique letters that were sampled in `f`.

```{r}
unique(f)
```

7. Use a function in the `forcats` packages that shows the unique letters that were sampled in `f`. How is this function different from the base R function? 

```{r}
fct_unique(f)
```
`fct_unique` function is specified for the factor variable. It will sort the unique values according to the levels of factor.
