# Recitation 7

__Today you learned...__

* How to create nice summary tables with the _gtsummary_ package. 

Now we'll practice using the functions we learned in class.

Load any necessary packages in your first chunk of code.

```{r , include=FALSE}
library(tidyverse)
library(gtsummary)
```

Please work out the following exercises sequentially. There may be more than one way to code the exercises.

## Part I

1. Load in the built-in `trial` data set (in gtsummary). It contains data from 200 patients who received one of two types of chemotherapy (Drug A or Drug B). The outcomes are tumor response and death.

Apply `tbl_summary` to the trial data. Describe the results in a few sentences.

```{r}
trial %>% tbl_summary()
```

2. Use `tbl_summary` to summarize continuous variables of the trial data by treatment. Show up to two decimals.

```{r }
trial %>% 
  select(trt, age, marker, ttdeath) %>% 
  tbl_summary(by = trt, digits = all_continuous() ~ 2)
```

3. Now compare the two groups using the add_p() function. What do you see?

```{r }
trial %>% 
  select(trt, age, marker, ttdeath) %>% 
  tbl_summary(by = trt, digits = all_continuous() ~ 2) %>% 
  add_p()
```

4. Add a column with overall summary statistics (not separated by drug type).

```{r }
trial %>% 
  select(trt, age, marker, ttdeath) %>% 
  tbl_summary(by = trt, digits = all_continuous() ~ 2) %>% 
  add_p() %>%
  add_overall()
```

5. Add a column with N (or N missing) for each variable.

```{r }
trial %>% 
  select(trt, age, marker, ttdeath) %>% 
  tbl_summary(by = trt, digits = all_continuous() ~ 2) %>% 
  add_p() %>%
  add_overall() %>% 
  add_n()
```

6. Add labels for the summary statistics in each row rather than as a footnote. Also, change the statistics so that they show "Mean (SD)".

```{r}
trial %>% 
  select(trt, age, marker, ttdeath) %>% 
  tbl_summary(by = trt, 
              statistic = all_continuous() ~ "{mean} ({sd})",
              digits = all_continuous() ~ 2) %>% 
  add_p() %>%
  add_overall() %>% 
  add_n() %>% 
  add_stat_label()
```

## Part II

Formulate an interesting question using this data set and use any functions you've learned so far to answer it. If you can, create a table of your results using gtsummary. Some groups will present their questions and code!

```{r }
#Answering Kaylinn's Question
trial %>% 
  select(age, response, trt) %>% 
  mutate(response = ifelse(response == 1, "Tumor Observed", "No Tumor")) %>%
  tbl_summary(by = trt)
```
