# Recitation 6

__Today you learned...__

* Useful functions for exploratory data analysis (EDA) like _case_when()_, _group_by()_, and _summarize()_. 

Now we'll practice using the functions we learned in class.

Load any necessary packages in your first chunk of code.

```{r, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(tidyverse)
```

Please work out the following exercises sequentially. There may be more than one way to code the exercises.

## Part I

1. Import the Airbnb data set as `bnb_data` and apply `clean_names()`.

```{r}
bnb_data <- read_csv("dataset/nyc_airbnb.csv") %>% 
  janitor::clean_names()
```

2. Arrange the data (and save it) in decreasing order by price. 

```{r }
bnb_data <- bnb_data %>% arrange(desc(price))
```

3. Use `if_else()` to create a new variable called `price_bin` that equals 1 if the Airbnb is no less than average, and 0 if it is cheaper than average.

```{r p1q3}
bnb_data <- bnb_data %>% 
  mutate(price_bin = ifelse(price >= mean(price), 1, 0))
```

4. Use `case_when()` to do the same thing. 

```{r}
bnb_data <- bnb_data %>% 
  mutate(price_bin = case_when(price >= mean(price) ~ 1,
                               price < mean(price) ~ 0)
  )

# another version
bnb_data <- bnb_data %>% 
  mutate(price_bin = case_when(price > mean(price) ~ 1,
                               TRUE ~ 0)
  )
```

5. Find the mean, median, sd, min, and max of the price variable for each neighborhood (`neighborhood_group`). Use `group_by()` and `summarize()`.

```{r}
bnb_data %>% 
  group_by(neighbourhood_group) %>% 
  summarize(mean = mean(price),
            sd = sd(price),
            min = min(price),
            max = max(price))
```

6. How many listings contain the word "cozy" (or "Cozy" or "COZY") in their names? 

```{r}
bnb_data %>% 
  filter(str_detect(name, 'cozy|Cozy|COZY')) %>% 
  nrow()
```

There are 4234 listings that contain the word "cozy", "Cozy", or "COZY". 

7. For each neighborhood, find the average price of listings containing the word "cozy", "Cozy", or "COZY" in their names.

```{r }
bnb_data %>% 
  filter(str_detect(name, 'cozy|Cozy|COZY')) %>% 
  group_by(neighbourhood_group) %>% 
  summarize(mean = mean(price))
```

8. Show the name and number of listings of the 5 hosts with the most listings containing the words "cozy", "Cozy", or "COZY". 

```{r}
bnb_data %>% 
  filter(str_detect(name, 'cozy|Cozy|COZY')) %>% 
  arrange(desc(calculated_host_listings_count)) %>% 
  select(host_name, calculated_host_listings_count) %>% 
  unique() %>% 
  head(5)

# Danielle's Version
bnb_data %>% 
  filter(str_detect(name, "cozy|Cozy|COZY")) %>% 
  group_by(host_id) %>% 
  reframe(name = host_name,
          cozy_count = n()) %>% 
  distinct() %>% 
  arrange(desc(cozy_count)) %>% 
  head(5)

# Lidio's Version
bnb_data %>% 
  filter(str_detect(name, "cozy|Cozy|COZY")) %>% 
  group_by(host_id, host_name) %>% 
  count(host_id) %>% 
  arrange(desc(n)) %>% 
  head(5)
```

9. Use `str()` to see the data types of all the variables. Change the `minimum_nights` variable to a factor variable. 

```{r}
str(bnb_data)

bnb_data <- bnb_data %>% 
  mutate(minimum_nights = as_factor(minimum_nights))
```

10. Group by room type and count the number of Airbnbs in each level of `minimum_nights`. 

```{r}
bnb_data %>% 
  group_by(room_type) %>% 
  select(minimum_nights) %>% 
  count(minimum_nights)

bnb_data |>
  group_by(room_type, minimum_nights) |>
  summarise(
    count = n()
  )
```
