# Final Review

```{r, message=FALSE}
# data analysis
library(tidyverse) 
# string
library(stringr) 
# factor
library(forcats) 
```


## Data Loading
* `read_csv`: load `csv` data table locally
* `janitor::clean_names()`: tidy up the column names of dataframe, it converts names to snake_case, removes special characters, and handles duplicated names.

```{r review-load-clean, message=F}
airbnb = read.csv("./dataset/nyc_airbnb.csv") 
airbnb = janitor::clean_names(airbnb)
colnames(airbnb)
```


## Pipe Operator
The pipe operator (`%>%` or `|>`) allows for cleaner and more readable code by chaining together a sequence of operations. Instead of nesting functions, the pipe operator allows for a left-to-right flow of data transformations.
```{r review-pipe-operator, message=F}
airbnb = read_csv("./dataset/nyc_airbnb.csv") %>% janitor::clean_names()
airbnb = read_csv("./dataset/nyc_airbnb.csv") |> janitor::clean_names()
airbnb |> head(10)
```


## `tidyverse`
### `dplyr` functions

- `select()`: Selects columns from a data frame.
- `filter()`: Filters rows based on specified conditions.
- `mutate()`: Adds new variables or transforms existing ones.
- `arrange()`: Orders rows by specified variables.
- `rename()`: Renames columns in a data frame. 
```{r review-dplyr}
airbnb |> 
  filter(neighbourhood_group == "Queens") |> 
  select(name:price) |>
  mutate(price_level = if_else(price >= 500, "Expensive", "Not expensive")) |>
  arrange(desc(price)) |>
  rename(host = host_name) |> 
  head(10)
```


### pivots

- `pivot_longer()`: Converts data from wide to long format.
- `pivot_wider()`: Converts data from long to wide format.
```{r review-pivot-data, message=FALSE}
year <- paste(2001:2020)
team <- letters[1:length(year)]
num <- sample(c(1:100), size = length(year)^2, replace = TRUE)

yr_data <- data.frame(matrix(num, nrow = length(year), ncol = length(year))) |>
  set_names(year) |> 
  mutate(id = row_number())
yr_data
```
```{r review-pivot-longer}
yr_data_long = yr_data |> pivot_longer(cols = paste(2001:2020), 
                                       names_to = "year", 
                                       values_to = "points")
yr_data_long
```

```{r review-pivot-wider}
yr_data_long |>
  pivot_wider(names_from = year, values_from = points)
```




### joins

- `right_join()`: Keeps all rows from the right data frame and matching rows from the left.
- `left_join()`: Keeps all rows from the left data frame and matching rows from the right.
- `inner_join()`: Keeps only rows that have matching values in both data frames.
- `full_join()`: Keeps all rows from both data frames, filling in `NA` for missing matches.
```{r review-join}
data1 <- tibble::tibble(id = 2:3, name = c("John", "Doe"))
data2 <- tibble::tibble(id = 1:2, salary = c(50000, 60000))

# right_join
data1 |> right_join(data2, by = "id")

# left_join
data1 |> left_join(data2, by = "id")

# inner_join
data1 |> inner_join(data2, by = "id")

# full_join
data1 |> full_join(data2, by = "id")
```

## String and Factor
Usage of `stringr` and `forcats` package
```{r review-str}
str_data <- "Hello, World!"
str_length <- str_length(str_data)
str_upper <- str_to_upper(str_data)
```

```{r review-fct}
fct_data <- factor(c("apple", "banana", "cherry", "apple"))
fct_count <- fct_count(fct_data)
fct_relevel <- fct_relevel(fct_data, "banana", "apple")
```



## EDA
- `case_when`: A versatile function for creating new variables based on multiple conditions.
- `group_by`: Groups data by one or more variables for subsequent operations.
- `summarize`: Calculates summary statistics for grouped data.
```{r review-eda}
airbnb |>
  group_by(neighbourhood_group) |>
  summarise(count = n()) |>
  arrange(desc(count)) |>
  mutate(level = case_when(
    count > 10000 ~ "Large",
    count > 2000 ~ "Medium", 
    count > 500 ~ "Small",
    .default = "Tiny"
  ))
```


summary tables with the _gtsummary_ package
```{r review-gt}
airbnb |>
  select(c(neighbourhood_group, price)) |>
  gtsummary::tbl_summary()
```


## `ggplot`
```{r review-ggplot1}
airbnb |>
  mutate(borough = neighbourhood_group |> factor()) |>
  ggplot() +
  geom_boxplot(aes(x = borough, y = price, color = borough)) +
  theme_bw() + 
  labs(
    title = "Boxplot of Airbnb price in different borough"
  ) + 
  theme(plot.title = element_text(hjust = 0.5))
```



## Control Statement
* `if-else`
```{r review-ifelse}
# if-else
x <- 10
if (x > 5) {
  result <- "Greater than 5"
} else {
  result <- "5 or less"
}
```

* `for` loop
```{r review-for}
for (i in 1:5) {
  print(i)
}
```

* `while` loop
```{r review-while}
count <- 1
while (count <= 5) {
  print(count)
  count <- count + 1
}
```

## Function
```{r review-func}
item_checkout = function(price, n, discount = 1){
  return(price * discount * n)
}

fruit_names = c("Apple", "Banana", "Cherry", "Dragonfruit", "Fig", "Kiwi", "Strawberry")
fruit = tibble(
  name = fruit_names, 
  price = runif(length(name), 0, 50),
  discount = runif(length(name)),
  count = sample(1:100, length(name))
)

fruit |>
  mutate(payment = item_checkout(price, count, discount))
```



